Appendix S1: Supporting Tables and Figures

### 2021-12-24 ms/SI.Rmd

---
title: "Appendix S1: Supporting Tables and Figures"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta et al.---\textit{American Journal of Botany} 2021 -- Appendix S1} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: pdf
# nocite: | 
---

@@

**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan. Curve fit with the `r pack("iNEXT")` package [@Hsieh2016]. Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line). Point indicates observed taxonomic richness (`r observed_data$qD` taxa, 95% confidence interval `r rich_interval` taxa). Taxa include species, subspecies, and varieties, excluding hybrids. Sampling completeness is estimated to be `r est_completeness` (95% confidence interval `r completeness_interval`).

@@

**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy. (A), Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells). (B), Percent of grid cells with redundancy = 0 by grain size. (C), Distribution of redundancy values by grain size. There is a sudden improvement in redundancy values from 0.1 to 0.2 degree grid-cells, but less improvement as grain size is increased beyond 0.2 degrees.

@@

a <- plot_env("A", "Area", "lat_area", "km^2")
b <- plot_env("B", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
c <- plot_env("C", "Precipitation", "precip", "mm")
d <- plot_env("D", "Precip. season.", "precip_season" , "")
e <- plot_env("E", "Temperature", "temp", "°C", function(x) x/10)
f <- plot_env("F", "Temp. season.", "temp_season") +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

@@

**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan. (A), rolling mean area (km^2^) in 1° latitudinal bands.  (B), % apomictic taxa. (C), annual precipitation. (D), percipitation seasonality (coefficient of variation).  (E), mean annual temperature. (F), temperature seasonality (standard deviation × 100).

@@

**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa. Points at nodes indicate bootstrap support. Scalebar 100 my. Tree visualized with the `r pack("ggtree")` R package [@Yu2017].

@@

**`r s_figure("abundance")`**. Observed number of specimens (A) and sampling redundancy (B) per 0.2° grid cell in the ferns of Japan.

@@

**`r s_figure("raw-lat")`**. (A) Raw taxonomic diversity (B), phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude. For (B) and (C), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths. Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan. (A), Phylogenetic diversity (PD). (B), Relative phylogenetic diversity (RPD). (C), Functional diversity (FD). (D), Relative functional diversity (RFD). Trendlines in (A, C) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

**`r s_figure("k-plot")`**. Selection of *k* for bioregions. (A), taxonomic bioregions. (B), phylogenetic bioregions. Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).

@@

**`r s_figure("dendrogram")`**. Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category. (A), taxonomic bioregions. (B), phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan. (A), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan. (B), *Deparia* is an example of a genus with possible recent radiation in southern Honshu. (C)--(F) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.

@@

**`r s_figure("apo-genus")`**. Rate of apomixis in genera Japanese ferns with > 1 apomictic taxon. Numbers next to each bar show number of apomictic taxa out of total richness per genus. Excludes non-native taxa and nothotaxa.

@@

**`r s_figure("modelfit-not-ult")`**. Relationship between (A) phylogenetic diversity (PD) and (B) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree). Ribbon shows 95% confidence interval of model. Line fit with the focal predictor variable while averaging over other predictors. For RPD, the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*P*-value) calculated using a two-tailed test (see Materials and Methods).

@@

**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions arranged by mean annual temperature and annual precipitation. Points shown with partial transparency; darker areas indicate overlapping points. Bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.

@@

a <- make_protected_plot("A", "Richness", richness_obs_p_upper > 0.95)
b <- make_protected_plot("B", "PD", pd_obs_p_upper > 0.95)
c <- make_protected_plot("C", "FD", fd_obs_p_upper > 0.95)
d <- make_protected_plot("D", "PE", pe_obs_p_upper > 0.95) +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

@@

**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas. Grid-cells with black outlines indicate significantly high biodiversity. Biodiversity metrics include (A) taxon richness, (B) phylogenetic diversity (PD), (C) functional diversity (FD), and (D) phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).

@@

a <- make_deer_plot("A", "Richness", richness_obs_p_upper > 0.95)
b <- make_deer_plot("B", "PD", pd_obs_p_upper > 0.95)
c <- make_deer_plot("C", "FD", fd_obs_p_upper > 0.95)
d <- make_deer_plot("D", "PE", pe_obs_p_upper > 0.95) +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

@@

**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*). Grid-cells with black outlines indicate significantly high biodiversity. Biodiversity metrics include (A) taxon richness, (B) phylogenetic diversity (PD), (C) functional diversity (FD), and (D) phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods). Deer range data sources include 1978 survey, 2003 survey, and range estimated from model based on 2003 survey data. Deer ranges plotted in layers in ascending order (estimated, 2003, 1978) such that upper layers may obscure lower layers; however, deer ranges generally expanded from 1978 to estimated range, so very few sites are obscured.

>>>>

---
title: "Appendix S1: Supporting Tables and Figures"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
header-includes: 
  - \usepackage{float} # keep figures from floating away
  - \usepackage{booktabs} # need for kable tables
  - \usepackage{colortbl} # need for kable tables
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # keep figures from floating away
  - \usepackage{fancyhdr} # enable header
  - \pagestyle{fancy}  # enable header
  - \fancypagestyle{plain}{\pagestyle{fancy}} # show header on first page
  - \renewcommand{\headrulewidth}{0pt} # delete line beneath header
  - \fancyhead[RE,RO]{Nitta \textit{et al}. Spatial phy. of Japanese ferns Appendix S1} # add header on right side
  - \fancyhead[LO,LE]{} # suppress automatic section header on left side
  - \cfoot{\thepage} # footer
editor_options: 
  chunk_output_type: console
params:
  doc_type: pdf
# nocite: | 
---

@@

**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan fit with the `r pack("iNEXT")` package [@Hsieh2016]. Grey shading indicates 95% confidence interval (not visible where it does not extend beyond the fitted line). Point indicates observed taxonomic richness (`r observed_data$qD` taxa, 95% confidence interval `r rich_interval` taxa). Taxa include species, subspecies, and varieties, excluding hybrids. Sampling completeness is estimated to be `r est_completeness` (95% confidence interval `r completeness_interval`).

@@

**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy. (a), Percent of grid cells with redundancy > 0.5 by grain size (0.1, 0.2, 0.3, or 0.4 degree grid-cells). (b), Percent of grid cells with redundancy = 0 by grain size. (c), Distribution of redundancy values by grain size. There is a sudden improvement in redundancy values from 0.1 to 0.2 degree grid-cells, but less improvement as grain size is increased beyond 0.2 degrees.

@@

a <- plot_env("a", "Area", "lat_area", "km^2")
b <- plot_env("b", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
c <- plot_env("c", "Precipitation", "precip", "mm")
d <- plot_env("d", "Precip. season.", "precip_season" , "")
e <- plot_env("e", "Temperature", "temp", "°C", function(x) x/10)
f <- plot_env("f", "Temp. season.", "temp_season") +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

@@

**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan. (a), rolling mean area (km^2^) in 1° latitudinal bands.  (b), % apomictic taxa. (c), annual precipitation. (d), percipitation seasonality (coefficient of variation).  (e), mean annual temperature. (f), temperature seasonality (standard deviation × 100).

@@

**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding hybrids. Points at nodes indicate bootstrap support. Scalebar 100 my. Tree visualized with the `r pack("ggtree")` R package [@Yu2017].

@@

**`r s_figure("abundance")`**. Observed number of specimens (a) and sampling redundancy (b) per 0.2° grid cell in the ferns of Japan.

@@

**`r s_figure("raw-lat")`**. (a) Raw taxonomic diversity (b), phylogenetic diversity, and (c) functional diversity of the ferns of Japan plotted by latitude. For (b) and (c), raw branchlengths were transformed to relative values by dividing each branchlength by the sum of all branchlengths. Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan. (a), Phylogenetic diversity (PD). (b), Relative phylogenetic diversity (RPD). (c), Functional diversity (FD). (d), Relative functional diversity (RFD). Trendlines in (a, c) fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

**`r s_figure("k-plot")`**. Selection of *k* for bioregions. (a), taxonomic bioregions. (b), phylogenetic bioregions. Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).

@@

**`r s_figure("dendrogram")`**. Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category. (a), taxonomic bioregions. (b), phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan. (a), *Equisetum* is an example of a possible refugial lineage with greater species richness in northern Japan. (b), *Deparia* is an example of a genus with possible recent radiation in southern Honshu. (c)--(f) show genera with high rates of apomictic taxa (`r s_figure("apo-genus")`), which tend to be greatest in southern Honshu.

@@

**`r s_figure("apo-genus")`**. Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon. Numbers next to each bar show number of apomictic taxa out of total richness per genus.

@@

**`r s_figure("modelfit-not-ult")`**. Relationship between (a) phylogenetic diversity (PD) and (b) relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree). Ribbon shows 95% confidence interval of model. Line fit with the focal predictor variable while averaging over other predictors. For RPD, the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*P*-value) calculated using a two-tailed test (see Materials and Methods).

@@

**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in (a) taxonomic or (b) phylogenetic bioregions arranged by mean annual temperature and annual precipitation. Points shown with partial transparency; darker areas indicate overlapping points. Bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.

@@

a <- make_protected_plot("a", "Richness", richness_obs_p_upper > 0.95)
b <- make_protected_plot("b", "PD", pd_obs_p_upper > 0.95)
c <- make_protected_plot("c", "FD", fd_obs_p_upper > 0.95)
d <- make_protected_plot("d", "PE", pe_obs_p_upper > 0.95) +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

@@

**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas. Grid-cells with black outlines indicate significantly high biodiversity. Biodiversity metrics include (a) taxon richness, (b) phylogenetic diversity (PD), (c) functional diversity (FD), and (d) phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods).

@@

a <- make_deer_plot("a", "Richness", richness_obs_p_upper > 0.95)
b <- make_deer_plot("b", "PD", pd_obs_p_upper > 0.95)
c <- make_deer_plot("c", "FD", fd_obs_p_upper > 0.95)
d <- make_deer_plot("d", "PE", pe_obs_p_upper > 0.95) +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

@@

**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*). Grid-cells with black outlines indicate significantly high biodiversity. Biodiversity metrics include (a) taxon richness, (b) phylogenetic diversity (PD), (c) functional diversity (FD), and (d) phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid-cells in the top 5% considered highly diverse (see Materials and Methods). Deer range data sources include 1978 survey, 2003 survey, and range estimated from model based on 2003 survey data. Deer ranges plotted in layers in ascending order (estimated, 2003, 1978) such that upper layers may obscure lower layers; however, deer ranges generally expanded from 1978 to estimated range, so very few sites are obscured.
